Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-03-17 analysis2_manuscript.tex
more edits

\author[1,*]{Graham Gower}
\author[2*]{Nathaniel S. Pope}
\author[2*]{Murillo F. Rodrigues}
\author[2*]{Silas Tittes}
\author[3*]{Linh N. Tran}

\author[4]{Ornob Alam}
\author[5]{Maria Izabel A. Cavassim}
\author[6]{Peter D. Fields}
\author[7]{Benjamin C. Haller}
\author[8,9]{Xin Huang}
\author[10]{Ben Jeffrey}
\author[5]{Christopher C. Kyriazis}
\author[2]{Kevin Korfmann}
\author[2]{Jiseon Min}
\author[11]{Inés Rebollo}
\author[2]{Clara T. Rehmann}
\author[2]{Chris C. R. Smith}
\author[2]{Scott T. Small}
\author[12]{Georgia Tsambos}

\author[10]{Yan Wong}
\author[13]{Gregor Gorjanc}
\author[14]{Christian D. Huber}
\author[13]{Yu Zhang}

\author[3]{Ryan N. Gutenkunst}
\author[10]{Jerome Kelleher}
\author[5]{Kirk E. Lohmueller}
\author[15]{Aaron P. Ragsdale}
\author[16]{Daniel R. Schrider}
\author[17]{Ilan Gronau}
\author[2,18]{Peter L. Ralph}
\author[2]{Andrew D. Kern}

\affil[*]{\small{These authors contributed equally to the paper.}}
\affil[1]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
\affil[3]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
\affil[4]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
\affil[5]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
\affil[6]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
\affil[7]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
\affil[8]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
\affil[9]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
\affil[10]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
\affil[11]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
\affil[12]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
\affil[13]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
\affil[14]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
\affil[15]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
\affil[16]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
\affil[17]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
\affil[18]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}

@@

    \subsection*{Quality control and software development}
    Following the protocol established previously for other aspects of \stdpopsim{} \citep{adrion2020community},
    new DFEs are subjet to a rigorous quality control (QC) process,
    by which a separate individual re-implements the same DFE, and the two implementations
    are programatically tested for equality.
    The QC process identified a substantial bug in one DFE:
    the \textit{Drosophila melanogaster} DFE \texttt{LognormalPlusPositive\_R16} \citet{booker2021selective}.
    Due to an error in coding the log-normal distribution, deleterious selection coefficients
    were several orders of magnitude too large.
    The bug affected results in a separate project from one of our groups
    that made use of this DFE before QC had been done.
    All changes and additions to the code must pass a comprehensive set of unit tests:
    at the time of writing, the library includes over 11,000 lines of testing code,
    that run 2,862 unit tests, including comparisons to QC implementations
    and cross-checks for agreement, including between \slim and \msprime implementations of the same models.
    For instance, for testing we make use of \slim's scriptability to separately record in-simulation metadata
    information about current population sizes, DFEs, etcetera,
    which are then programmatically compared to the models used to generate the \slim script
    that ran the simulation.
    We make use of automatic test ``code coverage'' tools to help ensure that all portions of the code base
    are validated by these unit tests.

@@

\section*{Acknowledgments}\label{acknowledgements}
We thank Tom Booker for discussions and encouragement to incorporate realistic BGS models into \stdpopsim.

>>>>

% author tranche explanation:
% co-first authors
\author[1,*]{Graham Gower}
\author[2,*]{Nathaniel S. Pope}
\author[2,*]{Murillo F. Rodrigues}
\author[2,*]{Silas Tittes}
\author[3,*]{Linh N. Tran}

% additional authors
\author[4]{Ornob Alam}
\author[5]{Maria Izabel A. Cavassim}
\author[6]{Peter D. Fields}
\author[7]{Benjamin C. Haller}
\author[8,9]{Xin Huang}
\author[10]{Ben Jeffrey}
\author[2]{Kevin Korfmann}
\author[5]{Christopher C. Kyriazis}
\author[2]{Jiseon Min}
\author[11]{Inés Rebollo}
\author[2]{Clara T. Rehmann}
\author[2]{Scott T. Small}
\author[2]{Chris C. R. Smith}
\author[12]{Georgia Tsambos}
\author[10]{Yan Wong}
\author[13]{Yu Zhang}

% PI additional authors
\author[14]{Christian D. Huber}
\author[13]{Gregor Gorjanc}
\author[15]{Aaron P. Ragsdale}

% PIs that were very involved in the project
\author[16]{Ilan Gronau}
\author[3]{Ryan N. Gutenkunst}
\author[10]{Jerome Kelleher}
\author[5]{Kirk E. Lohmueller}
\author[17]{Daniel R. Schrider}
\author[2,18]{Peter L. Ralph}
\author[2]{Andrew D. Kern}

\affil[*]{\small{These authors contributed equally to the paper.}}
\affil[1]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
\affil[3]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
\affil[4]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
\affil[5]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
\affil[6]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
\affil[7]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
\affil[8]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
\affil[9]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
\affil[10]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
\affil[11]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
\affil[12]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
\affil[13]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
\affil[14]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
\affil[15]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
\affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
\affil[17]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
\affil[18]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}

@@

    \subsection*{Quality control and software development}
    Following the protocol established previously for other aspects of \stdpopsim{} \citep{adrion2020community},
    new DFEs are subjet to a rigorous quality control (QC) process,
    by which a separate individual re-implements the same DFE, and the two implementations
    are programatically tested for equality.
    The QC process identified a substantial bug in one DFE:
    the \textit{Drosophila melanogaster} DFE \texttt{LognormalPlusPositive\_R16} \citet{booker2021selective}.
    Due to an error in coding the log-normal distribution, deleterious selection coefficients
    were several orders of magnitude too large.
    The bug affected results in a separate project from one of our groups
    that made use of this DFE before QC had been done.
    All changes and additions to the code must pass a comprehensive set of unit tests:
    at the time of writing, the library includes over 11,000 lines of testing code,
    that run 2,862 unit tests, including comparisons to QC implementations
    and cross-checks for agreement, including between \slim and \msprime implementations of the same models.
    For instance, for testing we make use of \slim's scriptability to separately record in metadata
    of the simulation current population sizes, DFEs, etcetera,
    which are then programmatically compared to the models used to generate the \slim script
    that ran the simulation.
    We make use of automatic test ``code coverage'' tools to help ensure that all portions of the code base
    are validated by these unit tests.

@@

\section*{Acknowledgments}\label{acknowledgements}
We thank Tom Booker for discussions and encouragement to incorporate realistic BGS models into \stdpopsim.
We thank the numerous users of the \stdpopsim package and well as
workshop participants who provided feedback and suggestions for improvements.
